---
title: "CUNY R Summer Bridge - Final Project"
author: "Team Triple Threat: Mariya, Josh, Bill"
date: "August 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset

First, let's read in the data from github and generate a DataFrame.

```{r }
# Reading data from our github, generating DataFrame

df <- read.csv("https://raw.githubusercontent.com/wcschuknecht/summer17/master/R/final_data.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
colnames(df) <- c("mpg", "cylinders", "displacement", "horsepower", "weight", "acceleration", "model_year", "origin", "car_name")
head(df)
```

## Data Exploration

Next, let's do some data exploration on our new DataFrame using various base methods on df.

```{r data exploration}

# Summary statistics to ascertain mean, median, min, max, and quartiles

summary(df)

# Structure informs the shape of the overall DataFrame and the data type of each variable

str(df)

```

## Data Wrangling
#### Rename the columns, add car make and engine size to the dataset, categorize misspellings, and classify models

```{r data wrangling}

names(df)[1] <- "MPG"
names(df)[2] <- "Cylinders"
names(df)[3] <- "Displacement"
names(df)[4] <- "Horsepower"
names(df)[5] <- "Weight"
names(df)[6] <- "Acceleration"
names(df)[7] <- "Year"
names(df)[8] <- "Origin"
names(df)[9] <- "Car_Name"
df$Car_Make <- sapply(strsplit(df$Car_Name," "), `[`, 1)
names(df)[10] <- "Car_Make"
df$Engine <- (df$Cylinders * df$Displacement) / 1000
names(df)[11] <- "Engine"

#
df[,10] = sapply(df[,10],as.character)
df$Car_Make[df$Car_Make == 'chevy'] <- "chevrolet"
df$Car_Make[df$Car_Make == 'chevroelt'] <- "chevrolet"
df$Car_Make[df$Car_Make == 'maxda'] <- "mazda"
df$Car_Make[df$Car_Make == 'mercedes-benz'] <- "mercedes"
df$Car_Make[df$Car_Make == 'toyouta'] <- "toyota"
df$Car_Make[df$Car_Make == 'vokswagen'] <- "volkswagen"
df$Car_Make[df$Car_Make == 'vw'] <- "volkswagen"
df$Car_Make[df$Car_Make == 'mercury'] <- "ford"
df$Car_Make[df$Car_Make == 'datsun'] <- "nissan"

head(df)

```

## Graphics
#### Scatterplot showing horsepower and MPG, colored by model year
#### Scatterplot featuring car make and MPG, colored by model year
#### Histogram displaying frequency of each MPG in ranges of 5
#### Box plot showing median of MPG across all cars in the data set

```{r graphics}

require(ggplot2)
ggplot(df, aes(x = Horsepower, y = MPG)) + geom_point(aes(color=Year))

ggplot(df, aes(x = Car_Make, y = MPG)) + geom_point(aes(color=Year)) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(df, aes(x = MPG)) + geom_histogram(breaks=seq(0, 50, by=5))

ggplot(df, aes(y = MPG, x = 1)) + geom_boxplot()

```

## Questions & Analysis

Question 1:  Does an increase in Horsepower have an effect on MPG?
Answer 1:  Yes. The more horsepower a vehicle has, the less gas efficient it is.  

Question 2:  Does Year have an effect on MPG efficiency?
Answer 2:  Yes. Over time, newer models of vehicles gas efficient.

Question 3:  Which Car Make is consistently the most gas efficient?
Answer 3:  Nissan, Volkswagen, Honda